Evolution of Processes in Delivering Reliable Software in the European Scientific Arena

### 2018-06-14 paper-ieee-trans.tex
Address the largely developed/mature components.

The reminder of this paper is
organized as follows. Section~\ref{sec:ev} details the evolution of research on
software reliability in EC--projects. Section~\ref{sec:ntsr} provides
information about new research trends in software reliability based upon the
experiences from the recent {\sl INDIGO--DataCloud} project (INDIGO--DC). Conclusions are drawn in Section~\ref{sec:con}.

@@

\item \textit{Unit and Functional testing}.
Changes involving the addition of new functionalities, are required to be tested. \textit{Regression
testing} in this context is accomplished by enforcing the definition and periodic execution
of the tests, covering past fixed issues, to ensure that they are not reintroduced during the development
activities. \textit{Unit testing} completes the source code testing evaluation by focusing on the code's
internal design. The SQA criteria of {\sl INDIGO--DataCloud} set a recommended threshold of 70\% coverage
for unit testing. In Fig.~\ref{fig:fig_unittest}, unit testing coverage is compared for each software
component between the two major releases, denoting an incremental trend over the course of the project:
code coverage increased from an average value of 50.32\% to 63.34\%, outlined by the respective dates of
the project's first (INDIGO-1) and second (INDIGO-2) releases. By INDIGO-2, 53\% of the software components
(17 out of 32) were over the threshold of the aforementioned project's SQA code coverage recommendation
(70\%), while 75\% (24 out of 32) of the product stack exceeded 50\% coverage.

@@

>>>>

The reminder of this paper is
organized as follows. Section~\ref{sec:b} summaries the essential context of the research problem. Section~\ref{sec:ev} details the evolution of research on
software reliability in EC--projects. Section~\ref{sec:ntsr} provides
information about new research trends in software reliability based upon the
experiences from the recent {\sl INDIGO--DataCloud} project (INDIGO--DC). Conclusions are drawn in Section~\ref{sec:con}.

\section{Background}
\label{sec:b}

In this section, we detail the level of knowledge requested in delivering reliable software in order 
to understand the effort employed in the EC-funded software projects over time. We divide literature 
into four different domains that intersect this work: 1) software requirements, 2) software coding,
3) metrics and measurements, 4) testing.

\emph{Software requirements.} They are defined at various stages of a software development process
as a specification of what should be implemented. According to IEEE \cite{radatz}, requirement
analysis is a process for studying user needs to arrive at a definition of system, hardware or software
requirements. Characteristics of requirements that can assess software reliability are consistency and
correctness \cite{sommerville}: the former aims at providing predictable, maintainable and reliable
results, and the latter usually refers to the satisfaction of certain goals.

\emph{Software coding.} According to IEEE \cite{radatz}, coding is the process of expressing a computer
program in a programming language. To meet specific requirements, a program can be developed by using
high-level or low-level programming languages that envelope different characteristics and may affect the
result and quality of software. Furthermore, software is generally written by geographically distributed
teams whose developers can change over time, therefore it is extremely important for source code to be
understandable \cite{tashtoush}. There are various instruments that can be adopted during the software
development process to easily adapt software \cite{buse, baggen}: programming features that influence the
code readability and maintainability and standardized code that promotes the comparability of results.

\emph{Metrics and measurements.} According to IEEE \cite{radatz}, metrics is a function, with input as 
the software data, and output is a value which could decide on how the given attribute affects the software.
Metrics and measurements have an important role in the development process \cite{fenton, capers}. Metrics
are used to improve the ability to control, identify and measure the essential parameters of software 
during its development, leading to improve the results of the software product \cite{goodman}. With a correct
usage of metrics, it is possible to determine which parts of the software require to be changed or modified.
In the literature, different types of metrics exist to be able to provide information on various software
aspects, such as complexity \cite{mccabe} and object orientation \cite{chidamber, lorenz}. Often they have 
similar definitions, so as a consequence they can be implemented in different ways according to the
interpretation of their definition. This is the reason why it has been always a challenge to select the
proper software metric tools.

\emph{Testing.} According to IEEE \cite{radatz}, testing is a process of operating a system or component
under specified conditions, observing or recording the results, and making an evaluation of some aspect of
system or component. Testing is an important factor for software quality and reliability improvement. This
activity is quite complicated due to the variety of programming languages, operating systems and hardware
platforms that evolve over time. In the current literature, there are many types of software
testing methods \cite{myers} that aim at finding errors, bugs and faults during the software development
process, possibly in the early stages. Over time, testing metrics have been defined to assess software
quality \cite{athanasiou}: code coverage is the most frequently used metric for test software reliability
and there exist many tools for dynamic code coverage estimation according to the programming language
\cite{horgan}. A dedicated effort have to be taken into account in order to integrate testing solutions
during the software development process.

@@

\item \textit{Unit and Functional testing}.
Changes involving the addition of new functionalities, are required to be tested. \textit{Regression
testing} in this context is accomplished by enforcing the definition and periodic execution
of the tests, covering past fixed issues, to ensure that they are not reintroduced during the development
activities. \textit{Unit testing} completes the source code testing evaluation by focusing on the code's
internal design. The SQA criteria of {\sl INDIGO--DataCloud} set a recommended threshold of 70\% coverage
for unit testing. In Fig.~\ref{fig:fig_unittest}, unit testing coverage is compared for each software
component between the two major releases, denoting an incremental trend over the course of the project:
code coverage increased from an average value of 50.32\% to 63.34\%, outlined by the respective dates of
the project's first (INDIGO--1) and second (INDIGO--2) releases. By INDIGO--2, 53\% of the software components
(17 out of 32) were over the threshold of the aforementioned project's SQA code coverage recommendation
(70\%), while 75\% (24 out of 32) of the product stack exceeded 50\% coverage. 
The indiviual records followed also a growing trend (only 6 out of 32 components had lower values in 
INDIGO--2). It is worth noting that Fig.~\ref{fig:fig_unittest} not only includes new software developments 
implemented from scratch within the project. Software tools and libraries from 
external open--source projects -- such as {\sl tosca-parser} or {\sl heat-translator} --, contributed upstream
by the project, and well--established products, involved in {\sl INDIGO--DataCloud}
but not contributing with the 100\% of their codebases -- such as {\sl dCache} --, are also considered in the
analysis. However the values given apply to the entire codebase, even for the latter case of products. These 
products challenged the application of the SQA policies described in this section, by two means:
\begin{itemize}
    \item Previously developed code was not refactored since it was not \textit{owned} by the project.
    \item An agreement on prevailing SQA policies for such cases where quality practices were already in place.
\end{itemize}

@@

\bibitem{radatz}
J. Radatz and A. Geraci and F. Katki.
IEEE Standard glossary of softare engineering terminology.
\emph{IEEE Std. 610.12-1990}, 1990.

\bibitem{sommerville}
I. Sommerville.
Integrated requiremetns envineering: A tutorial.
\emph{IEEE Software},
vol. 14, issue 1, pp. 16--23, 2005.

\bibitem{tashtoush}
Y. Tashtoush and Z. Odat, et al.
Impact of programming features on code readability.
\emph{International Journal of Software Engineering and its Applications}, 
vol. 7, issue 6, pp. 441--458, 2013.

\bibitem{buse}
R. P. Buse and W. R. Weimer.
Learning a metric for code readability.
\emph{IEEE Transactions on Software Engineering},
vol. 36, issue 4, pp. 546--558, 2010.

\bibitem{baggen}
R. Baggen and J. P. Correia, et al.
Standardized code quality benchmarking for improving software maintainability.
\emph{Software Quality Journal},
vol. 20, issue 2, pp. 287--307, 2012.

\bibitem{fenton}
N.  Fenton,  and  S.  L.  Pfleeger.
Software  Metrics:  A  Rigorous  and  Practical Approach,
\emph{International Thomson Computer
Press}, London, UK, second edition, 1997.

\bibitem{capers}
J. T. Capers.
Applied Software Measurement
\emph{McGrawHill}, 1991.

\bibitem{goodman}
P. Goodman.
Practical Implementation of Software Metrics, 
\emph{McGraw Hill}, London, 1993.

\bibitem{mccabe}
T. J. McCabe.
A   Complexity   Measure,
\emph{IEEE   Transactions   on   Software Engineering},
Vol. 2, No. 4, pp. 308-320, 1976.

\bibitem{chidamber}
S.  R.  Chidamber  and  C.  E  Kemerer.
A  metrics  suite  for  Object--oriented  design,
\emph{IEEE Trans. Software Engineering}, 
June 1994.

\bibitem{lorenz}
M. Lorenz and J. Kidd.
Object-Oriented Software Metrics: A Practical Approach,
\emph{IEEE publications}.

\bibitem{myers}
G. J. Myers and C. Sandler and T. Badgett.
The art of software testing.
\emph{John Wiley \& Sons}, 2011.

\bibitem{athanasiou}
D. Athanasiou and A. Nugroho and J. Visser and A. Zaidman.
Test code quality and its relation to issue handling performance.
\emph{IEEE Transactions on Software Engineering}, 
vol. 40, issue 7, pp. 1100--1125, 2014.

\bibitem{horgan}
J. R. Horgan and S. London and M. R. Lyu.
Achieving software quality with testing coverage measuring.
\emph{IEEE Computer},
vol. 27, issue 9, pp. 60--69, 1994.
